Kaggle-in-class Data Challenges Can Boost Student Learning

### 2018-05-07 paper-kaggle.Rmd
re-worked julias results

```{r Melb, fig.cap = "\\label{fig:MAST90083} Performance on regression and classification questions relative to total exam score (A, B) and overall question (C, D) for students by type of data competition in MAST90083. Differences in medians indicate improved performance.", fig.height=5, fig.width=6}
library(tidyverse)
library(gridExtra)
library(forcats)
UniMelbData <- read_csv(file="data/deidentified_UMelb.csv") 
UniMelbData <- UniMelbData %>% mutate(Proj_alloc=fct_recode(Proj_alloc, "regression"="Melbourne Price", "classification"="Spam classification"))
g_pte_class <- ggplot(UniMelbData, aes(Proj_alloc, pte_class)) +
  geom_boxplot() + 
  xlab("") + 
  ylab("CE") + 
  annotate("text", x=0.5, y=0.2, label="A") + 
  ggtitle("Classification Performance")
g_pte_reg <- ggplot(UniMelbData, aes(Proj_alloc, pte_reg)) + 
  geom_boxplot() + 
  xlab("") + 
  ylab("RE") + 
  annotate("text", x=0.5, y=0.4, label="B") + 
  ggtitle("Regression Performance")
g_ptq_class <- ggplot(UniMelbData, aes(Proj_alloc, ptq_class)) +
  geom_boxplot() + 
  xlab("Competition") + 
  ylab("CQ") + 
  annotate("text", x=0.5, y=1.0, label="C")
g_ptq_reg <- ggplot(UniMelbData, aes(Proj_alloc, ptq_reg)) + 
  geom_boxplot() + 
  xlab("Competition") + 
  ylab("RQ") + 
  annotate("text", x=0.5, y=1.0, label="D")
grid.arrange(g_pte_class, g_pte_reg, g_ptq_class, g_ptq_reg, ncol=2)
set.seed(88888)
pdif_C <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, pte_class)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_C <- c(pdif_C,
            median(x$pte_class[x$class=="classification"]-
                   median(x$pte_class[x$class=="regression"])))
}
dif_C <- median(UniMelbData$pte_class[UniMelbData$Proj_alloc=="classification"]-
       median(UniMelbData$pte_class[UniMelbData$Proj_alloc=="regression"]))
pdif_R <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, pte_reg)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_R <- c(pdif_R,
            median(x$pte_reg[x$class=="regression"]-                   median(x$pte_reg[x$class=="classification"])))
}
dif_R <- median(UniMelbData$pte_reg[UniMelbData$Proj_alloc=="regression"]-
       median(UniMelbData$pte_reg[UniMelbData$Proj_alloc=="classification"]))
pdif_CQ <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, ptq_class)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_CQ <- c(pdif_CQ,
            median(x$ptq_class[x$class=="classification"]-
            median(x$ptq_class[x$class=="regression"])))
}
dif_CQ <- median(UniMelbData$ptq_class[UniMelbData$Proj_alloc=="classification"]-
       median(UniMelbData$ptq_class[UniMelbData$Proj_alloc=="regression"]))
pdif_RQ <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, ptq_reg)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_RQ <- c(pdif_RQ,
            median(x$ptq_reg[x$class=="regression"]-
              median(x$ptq_reg[x$class=="classification"])))
}
dif_RQ <- median(UniMelbData$ptq_reg[UniMelbData$Proj_alloc=="regression"]-
       median(UniMelbData$ptq_reg[UniMelbData$Proj_alloc=="classification"]))
```

Table \ref{tab:Melb_Perm} shows the results of permutation testing of median difference between the groups. Generally the results support the competition improved performance. Students who participated in the kaggle challenge for classification scored `r round(dif_C, 2)` higher than those that did the regression competition, on the classification problem. Using a permutation test, this corresponds to a significant difference in medians, with $p$-value of `r round(length(pdif_C[pdif_C >= dif_C])/1000, 3)`.

```{r, caption = "\\label{tab:Melb_Perm} Permutation test $p$-values for results from MAST90083, for each of the different normalized scores."}
library(kableExtra)
pvals <- data.frame(cl = c("CE", "RE", "CQ", "RQ"),
  p=c(round(length(pdif_C[pdif_C >= dif_C])/1000, 3),
      round(length(pdif_R[pdif_R >= dif_R])/1000, 3),
      round(length(pdif_CQ[pdif_CQ >= dif_CQ])/1000, 3),
      round(length(pdif_RQ[pdif_RQ >= dif_RQ])/1000, 3)))
kable(t(pvals), row.names=FALSE)
```

@@

scores <- bind_rows(scores_5242, scores_2420)
ggplot(scores, aes(x=class, y=perf)) + 
  geom_boxplot() + 
  xlab("Participated in competition") +
  ylab("RE")
```

>>>>

```{r Melb, fig.cap = "\\label{fig:MAST90083} Performance on regression and classification questions relative to total exam score (A, B) and overall question (C, D) for students by type of data competition in MAST90083. Differences in medians indicate improved performance.", fig.height=3, fig.width=6}
library(tidyverse)
library(gridExtra)
library(forcats)
UniMelbData <- read_csv(file="data/deidentified_UMelb.csv") 
UniMelbData <- UniMelbData %>% mutate(Proj_alloc=fct_recode(Proj_alloc, "regression"="Melbourne Price", "classification"="Spam classification"))
UniMelbData <- UniMelbData %>% 
  mutate(perf_class = ((`Q1_p2`+`Q10_p3_A`+`Q10_p7_B`)/12)/(`Total_100`/100)) %>%
  mutate(perf_reg = ((`Q6_p2_A`+`Q6_p2_B`+`Q6_p2_C`+`Q8_p2_C`+`Q8_p2_D`+`Q16_p2_B`)/12)/(`Total_100`/100))

p1 <- ggplot(UniMelbData, aes(x=Proj_alloc, y=perf_class)) +
  geom_boxplot() +
  xlab("Classification competition") +
  ylab("Performance")
p2 <- ggplot(UniMelbData, aes(x=Proj_alloc, y=perf_reg)) +
  geom_boxplot() +
  xlab("Regression competition") +
  ylab("Performance")
grid.arrange(p1, p2, ncol=2)
```

```{r eval=FALSE}
# Didn't take into account difficulty and discrimination
g_pte_class <- ggplot(UniMelbData, aes(Proj_alloc, pte_class)) +
  geom_boxplot() + 
  xlab("") + 
  ylab("CE") + 
  annotate("text", x=0.5, y=0.2, label="A") + 
  ggtitle("Classification Performance")
g_pte_reg <- ggplot(UniMelbData, aes(Proj_alloc, pte_reg)) + 
  geom_boxplot() + 
  xlab("") + 
  ylab("RE") + 
  annotate("text", x=0.5, y=0.4, label="B") + 
  ggtitle("Regression Performance")
g_ptq_class <- ggplot(UniMelbData, aes(Proj_alloc, ptq_class)) +
  geom_boxplot() + 
  xlab("Competition") + 
  ylab("CQ") + 
  annotate("text", x=0.5, y=1.0, label="C")
g_ptq_reg <- ggplot(UniMelbData, aes(Proj_alloc, ptq_reg)) + 
  geom_boxplot() + 
  xlab("Competition") + 
  ylab("RQ") + 
  annotate("text", x=0.5, y=1.0, label="D")
grid.arrange(g_pte_class, g_pte_reg, g_ptq_class, g_ptq_reg, ncol=2)
set.seed(88888)
pdif_C <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, pte_class)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_C <- c(pdif_C,
            median(x$pte_class[x$class=="classification"]-
                   median(x$pte_class[x$class=="regression"])))
}
dif_C <- median(UniMelbData$pte_class[UniMelbData$Proj_alloc=="classification"]-
       median(UniMelbData$pte_class[UniMelbData$Proj_alloc=="regression"]))
pdif_R <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, pte_reg)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_R <- c(pdif_R,
            median(x$pte_reg[x$class=="regression"]-                   median(x$pte_reg[x$class=="classification"])))
}
dif_R <- median(UniMelbData$pte_reg[UniMelbData$Proj_alloc=="regression"]-
       median(UniMelbData$pte_reg[UniMelbData$Proj_alloc=="classification"]))
pdif_CQ <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, ptq_class)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_CQ <- c(pdif_CQ,
            median(x$ptq_class[x$class=="classification"]-
            median(x$ptq_class[x$class=="regression"])))
}
dif_CQ <- median(UniMelbData$ptq_class[UniMelbData$Proj_alloc=="classification"]-
       median(UniMelbData$ptq_class[UniMelbData$Proj_alloc=="regression"]))
pdif_RQ <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, ptq_reg)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_RQ <- c(pdif_RQ,
            median(x$ptq_reg[x$class=="regression"]-
              median(x$ptq_reg[x$class=="classification"])))
}
dif_RQ <- median(UniMelbData$ptq_reg[UniMelbData$Proj_alloc=="regression"]-
       median(UniMelbData$ptq_reg[UniMelbData$Proj_alloc=="classification"]))
```

```{r eval=FALSE}
# This code was used to examine the scores across all questions, 
# and understand the variability among the groups.
# Difficulty and discrimination for each question is computed. 
UniMelbData_Qs <- UniMelbData %>% 
  select(ID,Q1_p2:Q16_p2_B) %>%
  gather(Q, score, -ID) %>%
  separate(Q, c("Q", "pts", "part")) %>%
  mutate(pts = as.numeric(sub("p", "", pts))) %>%
  mutate(pct = score/pts*100)
diff_discr <- UniMelbData_Qs %>% 
  group_by(Q, part) %>%
  summarise(m = mean(pct, na.rm=TRUE), s = sd(pct, na.rm=TRUE))
diff_discr %>% arrange(m) 
diff_discr %>% arrange(desc(s))
library(plotly)
ggplot(diff_discr, aes(x=m, y=s, label=paste(Q, part))) + geom_point()
ggplotly()
keep <- diff_discr %>% filter(m<71, s>20)
# 1,10, 16B, 6A,B,C, 8C,D, 2,3,4
```

```{r}
pdif_cl <- NULL
pdif_reg <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, perf_class)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_cl <- c(pdif_cl,
            median(x$perf_class[x$class=="classification"])-
              median(x$perf_class[x$class=="regression"]))
  pdif_reg <- c(pdif_reg,
            median(x$perf_reg[x$class=="regression"])-
              median(x$perf_reg[x$class=="classification"]))
}
pdif_cl_true <- median(UniMelbData$perf_class[UniMelbData$Proj_alloc=="classification"])-
              median(UniMelbData$perf_class[UniMelbData$Proj_alloc=="regression"])
pdif_reg_true <- median(UniMelbData$perf_reg[UniMelbData$Proj_alloc=="regression"])-
              median(UniMelbData$perf_reg[UniMelbData$Proj_alloc=="classification"])
pval_cl <- length(pdif_cl[pdif_cl > pdif_cl_true])/1000
pval_reg <- length(pdif_reg[pdif_reg > pdif_reg_true])/1000
```

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|r|r|}\hline
Question Set & Median difference & Permutation $p$-value \\\hline
Classification & 0.250 & 0.012\\
Regression & 0.104 & 0.00\\\hline
\end{tabular}
\caption{Comparison of median difference in performance by competition group.}
\end{center}
\label{tab:Melb_Perm}
\end{table}

Table \ref{tab:Melb_Perm} shows the results of permutation testing of median difference between the groups. Generally the results support the competition improved performance. Students who participated in the kaggle challenge for classification scored higher than those that did the regression competition, on the classification problem. Using a permutation test, this corresponds to a significant difference in medians. Similarly the results show that students who did the regression challenge, performed better on these exam questions.

@@

scores <- bind_rows(scores_5242, scores_2420)
ggplot(scores, aes(x=class, y=perf)) + 
  geom_boxplot() + 
  xlab("Participated in competition") +
  ylab("Performance")
```
